---
title: "Graph SD Analyses - Spring 2018"
output: html_notebook
---

Data analysis for Witt, J. K. (in press).  Graph construction: An empirical investigation on setting the range of the y-axis.  Meta-Psychology.

Includes analysis scripts for all 5 experiments.


```{r init}
require(lme4)
require(lmerTest)
require(BayesFactor)
require(reshape2)
require(lsr)
require(wesanderson)
condCol <- wes_palette("GrandBudapest1")[c(3,2,4)]
condPCH <- c(19, 15, 17)

options(scipen=999)

setwd("C:/Users/lroesele.IVV5NET/Claude_Code/ReproAI/Meta Psychology/Volume 3 (2019)/07_Witt_GraphConstruction/ReproAI/Witt_GraphConstruction_895/data")

```

#Experiment 1
```{r Exp1data}

dt <- read.csv("axisSize 1-24.csv",header=T)

  dt$respCentered <- dt$resp - 2.5
  dt$corr <- ifelse(dt$effectSize == 0, 1, NA)
  dt$corr[which(dt$effectSize == 1)] <- 1.5  #split difference because d = .1 is bigger than a null effect but smaller than a "small" effect
  dt$corr[which(dt$effectSize == 3)] <- 2
  dt$corr[which(dt$effectSize == 5)] <- 3
  dt$corr[which(dt$effectSize == 8)] <- 4
  dt$corrCentered <- dt$corr - 2.5

  ab <-  which(colnames(dt) == "axisRange")
  colnames(dt)[ab] <- "graphType"
  if (length(ab)>0) {
    dt <- dt[which(dt$Subject < 10),]  #get Ss from Exp 1 only
  #  dt <- dt[which(dt$Subject > 9),]
  }
    



```

Get a prelimnary look at the data

```{r Exp 1 prelimLook}

  interaction.plot(dt$effectSize,dt$graphType,dt$resp)
  interaction.plot(dt$corr,dt$graphType,dt$resp, col=condCol, lwd=3, lty=1,type="b",pch=condPCH)

```



```{r Exp1 indivLM}
subjs <- sort(unique(dt$Subject))
N <- length(subjs)

saveAll <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j),]
      sMod <- lm(resp ~ corrCentered, data=df)
      
      sa <- sa+1
      saveAll$subj[sa] <- subjs[i]
      saveAll$graphType[sa] <- df$graphType[1]
      saveAll$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll$coef[sa] <- summary(sMod)$coefficients[2]
    }
}

saveAll$bias <- (saveAll$intercept - 2.5) / 2.5 * 100  #put as a percentage


boxplot(saveAll$coef)
boxplot(saveAll$intercept)
boxplot(saveAll$coef ~ saveAll$graphType)
boxplot(saveAll$bias ~ saveAll$graphType)
plot(saveAll$subj,saveAll$coef,pch=19,col=saveAll$graphType)

saveAll <- saveAll[which(saveAll$subj != 1),]
saveAll <- saveAll[which(saveAll$subj != 8),]

allSubjOuts <- c(101, 108)

```


```{r Exp1 Means and Tests}

print("Means for coefficients")
a <- aggregate(coef ~ graphType, data=saveAll, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
print(a)

ss <- dcast(saveAll,subj ~ graphType, value.var = "coef")
colnames(ss) <- c("subj","full","sd","min")
ss$sdFullDiff <- ss$sd - ss$full
ss$sdMinDiff <- ss$sd - ss$min


print("SD vs Full graphs")
t.test(ss$sd,ss$full,paired=T)
a <- t.test(ss$sd,ss$full,paired=T)
ttestBF(ss$sd,ss$full,paired=T,rscale="medium")
sdVsFulldz <- a$statistic / sqrt(length(ss$subj))
cohensD(ss$sd,ss$full,method = "paired")
psych::cohen.d.ci(sdVsFulldz,n1=length(ss$subj))


print("SD vs Minimal graphs")
t.test(ss$min,ss$sd,paired=T)
a <- t.test(ss$min,ss$sd,paired=T)
sdVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$sd,paired=T,rscale="medium")
cohensD(ss$min,ss$sd,method = "paired")
psych::cohen.d.ci(sdVsSmalldz,n1=length(ss$subj))


print("Full vs Small graphs")
t.test(ss$min,ss$full,paired=T)
a <- t.test(ss$min,ss$full,paired=T)
fullVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$full,paired=T,rscale="medium")
cohensD(ss$min,ss$full,method = "paired")
psych::cohen.d.ci(fullVsSmalldz,n1=length(ss$subj))



# Explore biases

ss <- dcast(saveAll,subj ~ graphType, value.var = "bias")
colnames(ss) <- c("subj","full","sd","min")

print("---------------")
print("Explore Biases")
a <- aggregate(bias ~ graphType, data=saveAll, mean)
a1 <- aggregate(bias ~ graphType, data=saveAll, sd)
a$sd <- a1$bias
a$graphType <- c("full","sd","min")
print(a)

  print(t.test(ss$min))
  a <- t.test(ss$min)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$min))

    print(t.test(ss$sd))
  a <- t.test(ss$sd)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$sd))
  
    print(t.test(ss$full))
  a <- t.test(ss$full)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$full))

print(t.test(ss$full,ss$sd,paired=T))
print(t.test(ss$min,ss$sd,paired=T))

```


```{r Exp1 Plots}

# Slopes
mm <- dcast(saveAll, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")

mm2 <- aggregate(coef ~ graphType, data=mm, mean)
mm2b <- aggregate(coef ~ graphType, data=mm, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm$coef) - (mean(mm2b$coef)*2), mean(mm$coef) + (mean(mm2b$coef)*2))
plot(c(1,2,3),mm2$coef,pch=19,cex=2,xlim=c(.5,3.5),ylim=yLim,bty="l",xaxt="n",ylab="Slope",xlab="Graph Axis Condition")
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"))


```

```{r exp1RespPlots}

for (i in allSubjOuts) {
  dt <- dt[which(dt$Subject != (i-100)),]
}

myCol <- condCol
myTitle <- c("Full","Standardized","Minimal")

bmp("exp1resps.bmp",height = 1000, width=1000)
par(mar=c(6,6,6,1))
par(mfrow=c(2,2))

for (i in 1:3) { 
  dd <- dt[which(as.integer(dt$graphType) == i),]
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=as.integer(dd$graphType),col=condCol[i], xlab="Depicted Effect Size",ylab = "Response",yaxt="n",xaxt="n",bty="l", main=myTitle[i], cex.lab=2.5,cex.axis=2, cex.main = 3)
  axis(side=1,at=c(.1,.3,.5,.8),labels = c(.1,.3,.5,.8),cex.axis=2)
  axis(side=2,at=c(1,2,3,4),labels = c(1,2,3,4),cex.axis=2)
}
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=4,col="white", xlab="Depicted Effect Size",ylab = "Correct Response",xaxt="n",yaxt="n",bty="l", main="Correct", cex.lab=2.5,cex.axis=2, cex.main = 3)
  points(jitter(dd$effectSize/10),jitter(dd$corr))
  axis(side=1,at=c(.1,.3,.5,.8),labels = c(.1,.3,.5,.8),cex.axis=2)
  axis(side=2,at=c(1,2,3,4),labels = c(1,2,3,4),cex.axis=2)

  dev.off()

print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==1 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==1)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp == 4)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp > 2)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))
  

```

#### Experiment 2

```{r Exp2 data}

dt <- read.csv("axisSize 1-24.csv")


  dt$corr <- ifelse(dt$effectSize == 0, 1, NA)
  dt$corr[which(dt$effectSize == 1)] <- 1.5
  dt$corr[which(dt$effectSize == 3)] <- 2
  dt$corr[which(dt$effectSize == 5)] <- 3
  dt$corr[which(dt$effectSize == 8)] <- 4
  dt$corrCentered <- dt$corr - 2.5


  ab <-  which(colnames(dt) == "axisRange")
  colnames(dt)[ab] <- "graphType"
  if (length(ab)>0) {
    dt <- dt[which(dt$Subject > 9),]#get Ss from Exp 2 only
  }
    



```

Get a prelimnary look at the data

```{r Exp 2 prelimLook}

  interaction.plot(dt$effectSize,dt$graphType,dt$resp)
  interaction.plot(dt$corr,dt$graphType,dt$resp)

```



```{r Exp2 indivLM}
subjs <- sort(unique(dt$Subject))
N <- length(subjs)

saveAll <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j),]
      sMod <- lm(resp ~ corrCentered, data=df)
      
      sa <- sa+1
      saveAll$subj[sa] <- subjs[i]
      saveAll$graphType[sa] <- df$graphType[1]
      saveAll$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll$coef[sa] <- summary(sMod)$coefficients[2]
    }
}

saveAll$bias <- (saveAll$intercept - 2.5) / 2.5 * 100  #put as a percentage


boxplot(saveAll$coef)
boxplot(saveAll$intercept)
boxplot(saveAll$coef ~ saveAll$graphType)
plot(saveAll$subj,saveAll$coef,pch=19,col=saveAll$graphType)

saveAll <- saveAll[which(saveAll$subj != 13),]
saveAll <- saveAll[which(saveAll$subj != 17),]
saveAll <- saveAll[which(saveAll$subj != 24),]

allSubjOuts <- c(allSubjOuts,213,217,224)

```


```{r Exp2 Means and Tests}

a <- aggregate(coef ~ graphType, data=saveAll, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
print(a)

ss <- dcast(saveAll,subj ~ graphType, value.var = "coef")
colnames(ss) <- c("subj","full","sd","min")
ss$sdFullDiff <- ss$sd - ss$full
ss$sdMinDiff <- ss$sd - ss$min


print("SD vs Full graphs")
t.test(ss$sd,ss$full,paired=T)
a <- t.test(ss$sd,ss$full,paired=T)
ttestBF(ss$sd,ss$full,paired=T,rscale="medium")
sdVsFulldz <- a$statistic / sqrt(length(ss$subj))
cohensD(ss$sd,ss$full,method = "paired")
psych::cohen.d.ci(sdVsFulldz,n1=length(ss$subj))


print("SD vs Minimal graphs")
t.test(ss$min,ss$sd,paired=T)
a <- t.test(ss$min,ss$sd,paired=T)
sdVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$sd,paired=T,rscale="medium")
cohensD(ss$min,ss$sd,method = "paired")
psych::cohen.d.ci(sdVsSmalldz,n1=length(ss$subj))


print("Full vs Small graphs")
t.test(ss$min,ss$full,paired=T)
a <- t.test(ss$min,ss$full,paired=T)
fullVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$full,paired=T,rscale="medium")
cohensD(ss$min,ss$full,method = "paired")
psych::cohen.d.ci(fullVsSmalldz,n1=length(ss$subj))



# Explore biases

ss <- dcast(saveAll,subj ~ graphType, value.var = "bias")
colnames(ss) <- c("subj","full","sd","min")

print("---------------")
print("Explore Biases")
a <- aggregate(bias ~ graphType, data=saveAll, mean)
a1 <- aggregate(bias ~ graphType, data=saveAll, sd)
a$sd <- a1$bias
a$graphType <- c("full","sd","min")
print(a)

  print(t.test(ss$min))
  a <- t.test(ss$min)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$min))

    print(t.test(ss$sd))
  a <- t.test(ss$sd)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$sd))
  
    print(t.test(ss$full))
  a <- t.test(ss$full)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$full))

print(t.test(ss$full,ss$sd,paired=T))
print(t.test(ss$min,ss$sd,paired=T))


```


```{r Exp2 Plots}

# Slopes
mm <- dcast(saveAll, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")

mm2 <- aggregate(coef ~ graphType, data=mm, mean)
mm2b <- aggregate(coef ~ graphType, data=mm, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm$coef) - (mean(mm2b$coef)*2), mean(mm$coef) + (mean(mm2b$coef)*2))
plot(c(1,2,3),mm2$coef,pch=19,cex=2,xlim=c(.5,3.5),ylim=yLim,bty="l",xaxt="n",ylab="Slope",xlab="Graph Axis Condition")
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"))



```


```{r exp2RespPlots}

myCol <- c("green","blue","red")
myTitle <- c("Full","Standardize","Minimal")
for (i in 1:3) { 
  dd <- dt[which(as.integer(dt$graphType) == i),]
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=as.integer(dd$graphType),col=myCol[i], xlab="Effect Size",ylab = "Response",yaxt="n",bty="l", main=myTitle[i])
  axis(side=2,at=seq(1,4),labels = seq(1,4))
}

print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==1 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==1)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp == 4)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp > 2)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))
  

```


#### Experiment 3
```{r Exp3 data}


dt <- read.csv("axisRangeEBv2 1-14.csv")


  dt$corr <- ifelse(dt$effectSize == 0, 1, NA)
  dt$corr[which(dt$effectSize == 1)] <- 1.5  #not sure what to do with this.  maybe exclude these experiments.
  dt$corr[which(dt$effectSize == 3)] <- 2
  dt$corr[which(dt$effectSize == 5)] <- 3
  dt$corr[which(dt$effectSize == 8)] <- 4
  dt$corrCentered <- dt$corr - 2.5


```

Get a prelimnary look at the data

```{r Exp 3 prelimLook}

  interaction.plot(dt$effectSize,dt$graphType,dt$resp)
  interaction.plot(dt$corr,dt$graphType,dt$resp)

```



```{r Exp3 indivLM}
subjs <- sort(unique(dt$Subject))
N <- length(subjs)

saveAll <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j),]
      sMod <- lm(resp ~ corrCentered, data=df)
      
      sa <- sa+1
      saveAll$subj[sa] <- subjs[i]
      saveAll$graphType[sa] <- df$graphType[1]
      saveAll$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll$coef[sa] <- summary(sMod)$coefficients[2]
    }
}

saveAll$bias <- (saveAll$intercept - 2.5) / 2.5 * 100  #put as a percentage

saveAll2 <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll2) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize > 1),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      sa <- sa+1
      saveAll2$subj[sa] <- subjs[i]
      saveAll2$graphType[sa] <- df$graphType[1]
      saveAll2$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll2$coef[sa] <- summary(sMod)$coefficients[2]
    }
}

saveAll3 <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll3) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize < 5),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      sa <- sa+1
      saveAll3$subj[sa] <- subjs[i]
      saveAll3$graphType[sa] <- df$graphType[1]
      saveAll3$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll3$coef[sa] <- summary(sMod)$coefficients[2]
    }
}
boxplot(saveAll$coef)
boxplot(saveAll$intercept)
boxplot(saveAll$coef ~ saveAll$graphType)
boxplot(saveAll2$coef ~ saveAll2$graphType)
boxplot(saveAll3$coef ~ saveAll3$graphType)
plot(saveAll$subj,saveAll$coef,pch=19,col=saveAll$graphType)

saveAll <- saveAll[which(saveAll$subj != 3),]
saveAll <- saveAll[which(saveAll$subj != 4),]

saveAll2 <- saveAll2[which(saveAll2$subj != 3),]
saveAll2 <- saveAll2[which(saveAll2$subj != 4),]
saveAll2 <- saveAll2[which(saveAll2$subj != 2),]
saveAll2 <- saveAll2[which(saveAll2$subj != 5),]

saveAll3 <- saveAll3[which(saveAll3$subj != 3),]
saveAll3 <- saveAll3[which(saveAll3$subj != 4),]
saveAll3 <- saveAll3[which(saveAll3$subj != 2),]
saveAll3 <- saveAll3[which(saveAll3$subj != 5),]

allSubjOuts <- c(allSubjOuts,303, 304, 305, 302)

```


```{r Exp3 Means and Tests}

#Print Means and SDs
a <- aggregate(coef ~ graphType, data=saveAll, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "allDs"
print(a)

a <- aggregate(coef ~ graphType, data=saveAll2, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll2, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "effMagnitude"
print(a)

a <- aggregate(coef ~ graphType, data=saveAll3, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll3, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "effVSnoEff"
print(a)



#Do analyses on slope coefficients

ss <- dcast(saveAll,subj ~ graphType, value.var = "coef")
colnames(ss) <- c("subj","full","sd","min")
ss$sdFullDiff <- ss$sd - ss$full
ss$sdMinDiff <- ss$sd - ss$min

ss2 <- dcast(saveAll2,subj ~ graphType, value.var = "coef")
colnames(ss2) <- c("subj","full","sd","min")
ss2$sdFullDiff <- ss2$sd - ss2$full
ss2$sdMinDiff <- ss2$sd - ss2$min

ss3 <- dcast(saveAll3,subj ~ graphType, value.var = "coef")
colnames(ss3) <- c("subj","full","sd","min")
ss3$sdFullDiff <- ss3$sd - ss3$full
ss3$sdMinDiff <- ss3$sd - ss3$min


print("Full vs Small graphs")
t.test(ss$min,ss$full,paired=T)
a <- t.test(ss$min,ss$full,paired=T)
fullVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$full,paired=T,rscale="medium")
cohensD(ss$min,ss$full,method = "paired")
psych::cohen.d.ci(fullVsSmalldz,n1=length(ss$subj))

dataOpts <- c("all","effOnly","no3")
comps <- c(1,3)
saveOuts <- as.data.frame(matrix(NA,ncol=13,nrow=3*2))
colnames(saveOuts) <- c("comp1","comp2","data","t","p","ll","dz","ul","BF","meanDiff","confLower","confUpper","df")
sa <- 0
currN <- length(ss$subj)
for (j in 1:length(comps)) {
    a <- t.test(ss$sd,ss[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[1]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]

        a <- t.test(ss2$sd,ss2[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[2]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]

        a <- t.test(ss3$sd,ss3[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[3]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]
    
}


print(saveOuts)


# Explore biases
ss <- dcast(saveAll,subj ~ graphType, value.var = "bias")
colnames(ss) <- c("subj","full","sd","min")

print("---------------")
print("Explore Biases")
print("---------------")
a <- aggregate(bias ~ graphType, data=saveAll, mean)
a1 <- aggregate(bias ~ graphType, data=saveAll, sd)
a$sd <- a1$bias
a$graphType <- c("full","sd","min")
print(a)

  print(t.test(ss$min))
  a <- t.test(ss$min)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$min))

    print(t.test(ss$sd))
  a <- t.test(ss$sd)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$sd))
  
    print(t.test(ss$full))
  a <- t.test(ss$full)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$full))

print(t.test(ss$full,ss$sd,paired=T))
print(t.test(ss$min,ss$sd,paired=T))


```


```{r Exp3 Plots}

# Slopes
mm <- dcast(saveAll, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")

mm2 <- aggregate(coef ~ graphType, data=mm, mean)
mm2b <- aggregate(coef ~ graphType, data=mm, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm$coef) - (mean(mm2b$coef)*2), mean(mm$coef) + (mean(mm2b$coef)*2))
plot(c(1,2,3),mm2$coef,pch=19,cex=2,xlim=c(.5,3.5),ylim=yLim,bty="l",xaxt="n",ylab="Slope",xlab="Graph Axis Condition")
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"))




```


```{r exp3RespPlots}

myCol <- c("green","blue","red")
myTitle <- c("Full","Standardize","Minimal")
for (i in 1:3) { 
  dd <- dt[which(as.integer(dt$graphType) == i),]
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=as.integer(dd$graphType),col=myCol[i], xlab="Effect Size",ylab = "Response",yaxt="n",bty="l", main=myTitle[i])
  axis(side=2,at=seq(1,4),labels = seq(1,4))
}

print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==1 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==1)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp == 4)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp > 2)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))
  

```


#### Experiment 4
```{r Exp4 data}


dt <- read.csv("axisRangeLineV2 1-20.csv")


  dt$corr <- ifelse(dt$effectSize == 0, 1, NA)
  dt$corr[which(dt$effectSize == 1)] <- 1.5  #not sure what to do with this.  maybe exclude these experiments.
  dt$corr[which(dt$effectSize == 3)] <- 2
  dt$corr[which(dt$effectSize == 5)] <- 3
  dt$corr[which(dt$effectSize == 8)] <- 4
  dt$corrCentered <- dt$corr - 2.5


```

Get a prelimnary look at the data

```{r Exp 4 prelimLook}

  interaction.plot(dt$effectSize,dt$graphType,dt$resp)
  interaction.plot(dt$corr,dt$graphType,dt$resp)

```



```{r Exp4 indivLM}
subjs <- sort(unique(dt$Subject))
N <- length(subjs)

saveAll <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j),]
      sMod <- lm(resp ~ corrCentered, data=df)
      
      sa <- sa+1
      saveAll$subj[sa] <- subjs[i]
      saveAll$graphType[sa] <- df$graphType[1]
      saveAll$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll$coef[sa] <- summary(sMod)$coefficients[2]
    }
}
saveAll$bias <- (saveAll$intercept - 2.5) / 2.5 * 100  #put as a percentage


saveAll2 <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll2) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize > 1),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      sa <- sa+1
      saveAll2$subj[sa] <- subjs[i]
      saveAll2$graphType[sa] <- df$graphType[1]
      saveAll2$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll2$coef[sa] <- summary(sMod)$coefficients[2]
    }
}

saveAll3 <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll3) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize < 5),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      sa <- sa+1
      saveAll3$subj[sa] <- subjs[i]
      saveAll3$graphType[sa] <- df$graphType[1]
      saveAll3$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll3$coef[sa] <- summary(sMod)$coefficients[2]
    }
}
boxplot(saveAll$coef)
boxplot(saveAll$intercept)
boxplot(saveAll$coef ~ saveAll$graphType)
boxplot(saveAll2$coef ~ saveAll2$graphType)
boxplot(saveAll3$coef ~ saveAll3$graphType)
plot(saveAll$subj,saveAll$coef,pch=19,col=saveAll$graphType)

subjOuts <- c(9,4,15,16)
for (i in 1:length(subjOuts)) {
  saveAll <- saveAll[which(saveAll$subj != subjOuts[i]),]
  saveAll2 <- saveAll2[which(saveAll2$subj != subjOuts[i]),]
  saveAll3 <- saveAll3[which(saveAll3$subj != subjOuts[i]),]
}

allSubjOuts <- c(allSubjOuts,409,404, 415,416)

```


```{r Exp4 Means and Tests}

#Print Means and SDs
a <- aggregate(coef ~ graphType, data=saveAll, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "allDs"
print(a)

a <- aggregate(coef ~ graphType, data=saveAll2, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll2, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "effMagnitude"
print(a)

a <- aggregate(coef ~ graphType, data=saveAll3, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll3, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "effVSnoEff"
print(a)



#Do analyses on slope coefficients

ss <- dcast(saveAll,subj ~ graphType, value.var = "coef")
colnames(ss) <- c("subj","full","sd","min")
ss$sdFullDiff <- ss$sd - ss$full
ss$sdMinDiff <- ss$sd - ss$min

ss2 <- dcast(saveAll2,subj ~ graphType, value.var = "coef")
colnames(ss2) <- c("subj","full","sd","min")
ss2$sdFullDiff <- ss2$sd - ss2$full
ss2$sdMinDiff <- ss2$sd - ss2$min

ss3 <- dcast(saveAll3,subj ~ graphType, value.var = "coef")
colnames(ss3) <- c("subj","full","sd","min")
ss3$sdFullDiff <- ss3$sd - ss3$full
ss3$sdMinDiff <- ss3$sd - ss3$min


print("Full vs Small graphs")
t.test(ss$min,ss$full,paired=T)
a <- t.test(ss$min,ss$full,paired=T)
fullVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$full,paired=T,rscale="medium")
cohensD(ss$min,ss$full,method = "paired")
psych::cohen.d.ci(fullVsSmalldz,n1=length(ss$subj))

dataOpts <- c("all","effOnly","no3")
comps <- c(1,3)
saveOuts <- as.data.frame(matrix(NA,ncol=13,nrow=3*2))
colnames(saveOuts) <- c("comp1","comp2","data","t","p","ll","dz","ul","BF","meanDiff","confLower","confUpper","df")
sa <- 0
currN <- length(ss$subj)
for (j in 1:length(comps)) {
    a <- t.test(ss$sd,ss[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[1]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]

        a <- t.test(ss2$sd,ss2[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[2]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]

        a <- t.test(ss3$sd,ss3[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[3]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]
    
}


print(saveOuts)


# Explore biases
ss <- dcast(saveAll,subj ~ graphType, value.var = "bias")
colnames(ss) <- c("subj","full","sd","min")

print("---------------")
print("Explore Biases")
print("---------------")
a <- aggregate(bias ~ graphType, data=saveAll, mean)
a1 <- aggregate(bias ~ graphType, data=saveAll, sd)
a$sd <- a1$bias
a$graphType <- c("full","sd","min")
print(a)

  print(t.test(ss$min))
  a <- t.test(ss$min)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$min))

    print(t.test(ss$sd))
  a <- t.test(ss$sd)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$sd))
  
    print(t.test(ss$full))
  a <- t.test(ss$full)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$full))

print(t.test(ss$full,ss$sd,paired=T))
print(t.test(ss$min,ss$sd,paired=T))


```


```{r Exp4 Plots}

# Slopes
mm <- dcast(saveAll, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")
mm$group <- 1

m4 <- mm

mm <- dcast(saveAll2, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")
mm$group <- 2

m4 <- rbind(m4,mm)

mm <- dcast(saveAll3, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")
mm$group <- 3

m4 <- rbind(m4,mm)


mm2 <- aggregate(coef ~ graphType + group, data=m4, mean)
mm2b <- aggregate(coef ~ graphType + group, data=m4, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm2$coef) - (mean(mm2b$coef)*2), mean(mm2$coef) + (mean(mm2b$coef)*2))
yLim <- c(min(mm2$coef) - .05,max(mm2$coef)+.05)
plot(as.integer(mm2$graphType),mm2$coef,pch=mm2$group,col=mm2$group,cex=2,xlim=c(.5,3.5),ylim=yLim,bty="l",xaxt="n",ylab="Slope",xlab="Graph Axis Condition")
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"))

plot(mm2$group,mm2$coef,pch=mm2$group,col=as.integer(mm2$graphType),cex=2,xlim=c(.5,3.5),ylim=yLim,bty="l",xaxt="n",ylab="Sensitivity",xlab="Graphs Included in Analysis")
axis(side=1,at=c(1,2,3),labels = c("All",".3 - .8","0 - .3"))





```



```{r exp4RespPlots}

myCol <- c("green","blue","red")
myTitle <- c("Full","Standardize","Minimal")
for (i in 1:3) { 
  dd <- dt[which(as.integer(dt$graphType) == i),]
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=as.integer(dd$graphType),col=myCol[i], xlab="Effect Size",ylab = "Response",yaxt="n",bty="l", main=myTitle[i])
  axis(side=2,at=seq(1,4),labels = seq(1,4))
}

print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==1 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==1)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp == 4)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp > 2)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))
  

```


#### Experiment 5
```{r Exp5 data}


dt <- read.csv("axisRangeLine 1-14.csv")


  dt$corr <- ifelse(dt$effectSize == 0, 1, NA)
  dt$corr[which(dt$effectSize == 1)] <- 1.5  #not sure what to do with this.  maybe exclude these experiments.
  dt$corr[which(dt$effectSize == 3)] <- 2
  dt$corr[which(dt$effectSize == 5)] <- 3
  dt$corr[which(dt$effectSize == 8)] <- 4
  dt$corrCentered <- dt$corr - 2.5


```

Get a prelimnary look at the data

```{r Exp 5 prelimLook}

  interaction.plot(dt$effectSize,dt$graphType,dt$resp)
  interaction.plot(dt$corr,dt$graphType,dt$resp)

```



```{r Exp5 indivLM}
subjs <- sort(unique(dt$Subject))
N <- length(subjs)

saveAll <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j),]
      sMod <- lm(resp ~ corrCentered, data=df)
      
      sa <- sa+1
      saveAll$subj[sa] <- subjs[i]
      saveAll$graphType[sa] <- df$graphType[1]
      saveAll$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll$coef[sa] <- summary(sMod)$coefficients[2]
    }
}
saveAll$bias <- (saveAll$intercept - 2.5) / 2.5 * 100  #put as a percentage


saveAll2 <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll2) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize > 1),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      sa <- sa+1
      saveAll2$subj[sa] <- subjs[i]
      saveAll2$graphType[sa] <- df$graphType[1]
      saveAll2$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll2$coef[sa] <- summary(sMod)$coefficients[2]
    }
}

saveAll3 <- as.data.frame(matrix(NA,ncol=4, nrow=N*3))
colnames(saveAll3) <- c("subj","graphType","intercept","coef")
sa <- 0

for (i in 1:N) {
  for (j in 1:3) {
      df <- dt[which(dt$Subject == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize < 5),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      sa <- sa+1
      saveAll3$subj[sa] <- subjs[i]
      saveAll3$graphType[sa] <- df$graphType[1]
      saveAll3$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll3$coef[sa] <- summary(sMod)$coefficients[2]
    }
}
boxplot(saveAll$coef)
boxplot(saveAll$intercept)
boxplot(saveAll$coef ~ saveAll$graphType)
boxplot(saveAll2$coef ~ saveAll2$graphType)
boxplot(saveAll3$coef ~ saveAll3$graphType)
plot(saveAll$subj,saveAll$coef,pch=19,col=saveAll$graphType)

subjOuts <- c(7, 13)  #11 is a maybe outlier but doesn't really count
for (i in 1:length(subjOuts)) {
  saveAll <- saveAll[which(saveAll$subj != subjOuts[i]),]
  saveAll2 <- saveAll2[which(saveAll2$subj != subjOuts[i]),]
  saveAll3 <- saveAll3[which(saveAll3$subj != subjOuts[i]),]
}

allSubjOuts <- c(allSubjOuts,507, 513)


```


```{r Exp5 Means and Tests}

#Print Means and SDs
a <- aggregate(coef ~ graphType, data=saveAll, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "allDs"
print(a)

a <- aggregate(coef ~ graphType, data=saveAll2, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll2, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "effMagnitude"
print(a)

a <- aggregate(coef ~ graphType, data=saveAll3, mean)
a1 <- aggregate(coef ~ graphType, data=saveAll3, sd)
a$sd <- a1$coef
a$graphType <- c("full","sd","min")
a$cond <- "effVSnoEff"
print(a)



#Do analyses on slope coefficients

ss <- dcast(saveAll,subj ~ graphType, value.var = "coef")
colnames(ss) <- c("subj","full","sd","min")
ss$sdFullDiff <- ss$sd - ss$full
ss$sdMinDiff <- ss$sd - ss$min

ss2 <- dcast(saveAll2,subj ~ graphType, value.var = "coef")
colnames(ss2) <- c("subj","full","sd","min")
ss2$sdFullDiff <- ss2$sd - ss2$full
ss2$sdMinDiff <- ss2$sd - ss2$min

ss3 <- dcast(saveAll3,subj ~ graphType, value.var = "coef")
colnames(ss3) <- c("subj","full","sd","min")
ss3$sdFullDiff <- ss3$sd - ss3$full
ss3$sdMinDiff <- ss3$sd - ss3$min


print("Full vs Small graphs")
t.test(ss$min,ss$full,paired=T)
a <- t.test(ss$min,ss$full,paired=T)
fullVsSmalldz <- a$statistic / sqrt(length(ss$subj))
ttestBF(ss$min,ss$full,paired=T,rscale="medium")
cohensD(ss$min,ss$full,method = "paired")
psych::cohen.d.ci(fullVsSmalldz,n1=length(ss$subj))

dataOpts <- c("all","effOnly","no3")
comps <- c(1,3)
saveOuts <- as.data.frame(matrix(NA,ncol=13,nrow=3*2))
colnames(saveOuts) <- c("comp1","comp2","data","t","p","ll","dz","ul","BF","meanDiff","confLower","confUpper","df")
sa <- 0
currN <- length(ss$subj)
for (j in 1:length(comps)) {
    a <- t.test(ss$sd,ss[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[1]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]

        a <- t.test(ss2$sd,ss2[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[2]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]

        a <- t.test(ss3$sd,ss3[,1+comps[j]],paired=T)
    
    sa <- sa+1
    saveOuts$comp1[sa] <- 2
    saveOuts$comp2[sa] <- comps[j]
    saveOuts$data[sa] <- dataOpts[3]
    saveOuts$t[sa] <- a$statistic
    saveOuts$p[sa] <- a$p.value
    saveOuts$df[sa] <- a$parameter
    dz <- a$statistic / sqrt(currN)
    saveOuts[sa,6:8] <- round(psych::cohen.d.ci(dz,n1=currN),2)
    saveOuts$BF[sa] <- exp(ttest.tstat(t=a$statistic, n1=currN, rscale = 0.707)[['bf']])
    saveOuts$meanDiff[sa] <- a$estimate
    saveOuts$confLower[sa] <- a$conf.int[1]
    saveOuts$confUpper[sa] <- a$conf.int[2]
    
}


print(saveOuts)


# Explore biases
ss <- dcast(saveAll,subj ~ graphType, value.var = "bias")
colnames(ss) <- c("subj","full","sd","min")

print("---------------")
print("Explore Biases")
print("---------------")
a <- aggregate(bias ~ graphType, data=saveAll, mean)
a1 <- aggregate(bias ~ graphType, data=saveAll, sd)
a$sd <- a1$bias
a$graphType <- c("full","sd","min")
print(a)

  print(t.test(ss$min))
  a <- t.test(ss$min)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$min))

    print(t.test(ss$sd))
  a <- t.test(ss$sd)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$sd))
  
    print(t.test(ss$full))
  a <- t.test(ss$full)
  psych::cohen.d.ci(a$statistic / sqrt(length(ss$subj)),n1=length(ss$subj))
  print(ttestBF(ss$full))

print(t.test(ss$full,ss$sd,paired=T))
print(t.test(ss$min,ss$sd,paired=T))



```


```{r Exp5 Plots}

# Slopes
mm <- dcast(saveAll, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")

mm2 <- aggregate(coef ~ graphType, data=mm, mean)
mm2b <- aggregate(coef ~ graphType, data=mm, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm$coef) - (mean(mm2b$coef)*2), mean(mm$coef) + (mean(mm2b$coef)*2))
plot(c(1,2,3),mm2$coef,pch=19,cex=2,xlim=c(.5,3.5),ylim=yLim,bty="l",xaxt="n",ylab="Slope",xlab="Graph Axis Condition")
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"))





```


```{r exp5RespPlots}

myCol <- c("green","blue","red")
myTitle <- c("Full","Standardize","Minimal")
for (i in 1:3) { 
  dd <- dt[which(as.integer(dt$graphType) == i),]
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=as.integer(dd$graphType),col=myCol[i], xlab="Effect Size",ylab = "Response",yaxt="n",bty="l", main=myTitle[i])
  axis(side=2,at=seq(1,4),labels = seq(1,4))
}

print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==1 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==1)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp == 4)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp > 2)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))
  

```

#### All Experiments Combined
```{r doAllExps}


dataSets <- c("axisSize 1-24.csv","axisRangeEBv2 1-14.csv","axisRangeLine 1-14.csv", "axisRangeLineV2 1-20.csv")
dataTitle <- c("Bar - 1SD","Bar - .6 SD w Error Bars","Line - .5 SD","Line - .7 SD")

  dd <- read.csv(dataSets[1],header=T)
  ab <-  which(colnames(dd) == "axisRange")
  colnames(dd)[ab] <- "graphType"
  dd$exp <- 1
  dd$exp[which(dd$Subject > 9)] <- 2
  dd$errBarType <- NA
  dd$cycle <- NA
  dd$sample <- NA
  dd$isBar <- 1
  dt <- dd
  
  dd <- read.csv(dataSets[2],header=T)
  dd$exp <- 3
  dd$Session <- 1
  dd$isBar <- 1
  dt <- rbind(dt,dd)
  
  dd <- read.csv(dataSets[3],header=T)
  dd <- dd[,c(1,3,4,6,7,8,9)]
  dd$exp <- 5
  dd$Session <- 1
  dd$graphImg <- NA
  dd$errBarType <- NA
  dd$isBar <- 0
  dt <- rbind(dt,dd)
  
  dd <- read.csv(dataSets[4],header=T)
  dd$exp <- 4
  dd$Session <- 1
  dd$errBarType <- NA
  dd$isBar <- 0
  dt <- rbind(dt,dd)
  
  dt$corr <- ifelse(dt$effectSize == 0, 1, NA)
  dt$corr[which(dt$effectSize == 1)] <- 1.5
  dt$corr[which(dt$effectSize == 3)] <- 2
  dt$corr[which(dt$effectSize == 5)] <- 3
  dt$corr[which(dt$effectSize == 8)] <- 4
  dt$corrCentered <- dt$corr - 2.5

  dt$graphType[which(as.character(dt$graphType) == "sd")] <- "SD"
  dt$graphType[which(as.character(dt$graphType) == "full")] <- "Full"
  dt$graphType[which(as.character(dt$graphType) == "small")] <- "Small"

  
  
  allSubjOuts <- c(101,108,213,217,224,303,304,305,302,409,404,415,416,507,513)

dt$subj2 <- dt$exp * 100 + dt$Subject
for (i in allSubjOuts) {
  dt <- dt[which(dt$subj2 != i),]
}
  
```


```{r analyzeAll}

subjs <- sort(unique(dt$subj2))
N <- length(subjs)

#For all trials
saveAll <- as.data.frame(matrix(NA,ncol=5, nrow=N*3))
colnames(saveAll) <- c("subj","graphType","intercept","coef","exp")
sa <- 0

#For trials with d > .10
saveAll2 <- as.data.frame(matrix(NA,ncol=5, nrow=N*3))
colnames(saveAll2) <- c("subj","graphType","intercept","coef","exp")


# Set value of minimum effect size to determine which data are included.
for (i in 1:N) {
  for (j in 4:6) {
      df <- dt[which(dt$subj2 == subjs[i] & as.integer(dt$graphType) == j),]
      sMod <- lm(resp ~ corrCentered, data=df)
      
      sa <- sa+1
      saveAll$subj[sa] <- subjs[i]
      saveAll$graphType[sa] <- df$graphType[1]
      saveAll$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll$coef[sa] <- summary(sMod)$coefficients[2]
      saveAll$exp[sa] <- df$exp[1]
      saveAll$bias[sa] <- (saveAll$intercept[sa] - 2.5) / 2.5 * 100  #put as a percentage

      df <- dt[which(dt$subj2 == subjs[i] & as.integer(dt$graphType) == j & dt$effectSize > 1),]
      sMod <- lm(resp ~ scale(corr,scale=F,center=T), data=df)
      
      saveAll2$subj[sa] <- subjs[i]
      saveAll2$graphType[sa] <- df$graphType[1]
      saveAll2$intercept[sa] <- summary(sMod)$coefficients[1]
      saveAll2$coef[sa] <- summary(sMod)$coefficients[2]
      saveAll2$exp[sa] <- df$exp[1]
      saveAll2$bias[sa] <- (saveAll2$intercept[sa] - 2.5) / 2.5 * 100  #put as a percentage

      
    }
}

saveAll$graphType[which(saveAll$graphType == 4)] <- 2 #SD
saveAll$graphType[which(saveAll$graphType == 5)] <- 1 #Full
saveAll$graphType[which(saveAll$graphType == 6)] <- 3 #Small

saveAll2$graphType[which(saveAll2$graphType == 4)] <- 2 #SD
saveAll2$graphType[which(saveAll2$graphType == 5)] <- 1 #Full
saveAll2$graphType[which(saveAll2$graphType == 6)] <- 3 #Small


```

```{r plotAllSlopes}

# Slopes
mm <- dcast(saveAll2, subj ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")

mm2 <- aggregate(coef ~ graphType, data=mm, mean)
mm2b <- aggregate(coef ~ graphType, data=mm, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm$coef) - (mean(mm2b$coef)*2), mean(mm$coef) + (mean(mm2b$coef)*2))  #within SDs
avgSD <- (sd(saveAll2$coef[which(saveAll2$graphType == 1)]) + sd(saveAll2$coef[which(saveAll2$graphType == 2)]) + sd(saveAll2$coef[which(saveAll2$graphType == 3)])) / 3
yLim2 <- c(mean(mm$coef) - (avgSD*1.5), mean(mm$coef) + (avgSD*1.5))
bmp("allSlopesByCond3.bmp",width=500,height=500)
  par(mar=c(5.1,5.1,2.1,.1))

plot(c(1,2,3),mm2$coef,pch=19,cex=3,xlim=c(.5,3.5),ylim=yLim2,bty="l",xaxt="n",ylab="Sensitivity",xlab="Graph Axis Condition",cex.lab=1.5,cex.axis=1.5)
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"),cex.axis=1.5)
for(i in 1:3) {
  segments(i,mm2$coef[i] - mm2$se[i],i,mm2$coef[i] + mm2$se[i],col="red",lwd=3)
}
dev.off()





```

```{r plotAllBiases}

# Intercepts
mm <- dcast(saveAll, subj ~ graphType, value.var = "bias")
colnames(mm) <- c("subj","full","sd","minimal")
mm$subjAvg <- apply(mm[,2:4],1,mean)
mm$adjustVal <- mean(mm$subjAvg) - mm$subjAvg
mm$fullW <- mm[,2] + mm$adjustVal
mm$sdW <- mm[,3] + mm$adjustVal
mm$minimalW <- mm[,4] + mm$adjustVal
print("check them")
for (i in 1:3) {
  print(paste(round(mean(mm[,i+1]),3),"vs",round(mean(mm[,i+6]),3)))
}
mm <- mm[,c(1,7,8,9)]
mm <- melt(mm,id="subj")
colnames(mm) <- c("subj","graphType","coef")

mm2 <- aggregate(coef ~ graphType, data=mm, mean)
mm2b <- aggregate(coef ~ graphType, data=mm, sd)
mm2$se <- mm2b$coef / sqrt(length(mm$subj))
yLim <- c(mean(mm$coef) - (mean(mm2b$coef)*3), mean(mm$coef) + (mean(mm2b$coef)*3))
avgSD <- (sd(saveAll$bias[which(saveAll2$graphType == 1)]) + sd(saveAll$bias[which(saveAll$graphType == 2)]) + sd(saveAll$bias[which(saveAll$graphType == 3)])) / 3
yLim2 <- c(mean(mm$coef) - (avgSD*2), mean(mm$coef) + (avgSD*2))
bmp("allBiasesByCond3.bmp",width=500,height=500)
  par(mar=c(5.1,5.1,2.1,.1))

plot(c(1,2,3),mm2$coef,pch=19,cex=3,xlim=c(.5,3.5),ylim=yLim2,bty="l",xaxt="n",ylab="Bias (%)",xlab="Graph Axis Condition",cex.lab=1.5,cex.axis=1.5)
axis(side=1,at=c(1,2,3),labels = c("Full","Standardized","Minimal"),cex.axis=1.5)
for(i in 1:3) {
  segments(i,mm2$coef[i] - mm2$se[i],i,mm2$coef[i] + mm2$se[i],col="red",lwd=3)
}
abline(h=0,lty=2,lwd=1)
dev.off()

```

```{r allPropResp}

print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==5 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==5)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==6 & dt$resp == 4 & dt$effectSize > 0)]) / length(dt$Subject[which(as.integer(dt$graphType)==6 & dt$effectSize > 0)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==6 & dt$resp > 2 & dt$effectSize > 0)]) / length(dt$Subject[which(as.integer(dt$graphType)==6 & dt$effectSize > 0)]))
  


```

```{r plotEachSubjByExp}

# Slopes
mm <- dcast(saveAll2, subj + exp ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","exp","full","sd","minimal")

bmp("SlopesByCondExp3.bmp",width=500,height=500)
  par(mar=c(5.1,5.1,2.1,.1))

plot(rep(.7,length(mm$subj))+(mm$exp/10),mm$full,xlim=c(.5,3.5),ylim=c(-.1,1.1),bty="l",ylab="Sensitivity",xlab="Graph Type",xaxt="n",col=rainbow(6)[mm$exp], pch=15, cex.lab=1.5, cex.axis=1.5, cex=2)
points(rep(1.7,length(mm$subj))+(mm$exp/10),mm$sd,col=rainbow(6)[mm$exp],pch=16, cex=2)
points(rep(2.7,length(mm$subj))+(mm$exp/10),mm$minimal,col=rainbow(6)[mm$exp],pch=17,cex=2)
axis(side=1,at=c(1,2,3),labels=c("Full","Standardized","Minimal"), cex.axis=1.5)
abline(h=0)
abline(h=1,lty=2)
dev.off()


mm <- dcast(saveAll, subj + exp ~ graphType, value.var = "coef")
colnames(mm) <- c("subj","exp","full","sd","minimal")

bmp("SlopesByCondExpAllTrials3.bmp",width=500,height=500)
  par(mar=c(5.1,5.1,2.1,.1))

plot(rep(.7,length(mm$subj))+(mm$exp/10),mm$full,xlim=c(.5,3.5),ylim=c(-.1,1.1),bty="l",ylab="Sensitivity",xlab="Graph Type",xaxt="n",col=rainbow(6)[mm$exp], pch=15, cex.lab=1.5, cex.axis=1.5, cex=2)
points(rep(1.7,length(mm$subj))+(mm$exp/10),mm$sd,col=rainbow(6)[mm$exp],pch=16, cex=2)
points(rep(2.7,length(mm$subj))+(mm$exp/10),mm$minimal,col=rainbow(6)[mm$exp],pch=17, cex=2)
axis(side=1,at=c(1,2,3),labels=c("Full","Standardized","Minimal"), cex.axis=1.5)
abline(h=0)
abline(h=1,lty=2)
dev.off()


mm <- dcast(saveAll, subj + exp ~ graphType, value.var = "bias")
colnames(mm) <- c("subj","exp","full","sd","minimal")

bmp("BiasByCondExpAllTrials3.bmp",width=500,height=500)
  par(mar=c(5.1,5.1,2.1,.1))

plot(rep(.7,length(mm$subj))+(mm$exp/10),mm$full,xlim=c(.5,3.5),ylim=c(-55,55),bty="l",ylab="Bias (%)",xlab="Graph Type",xaxt="n",col=rainbow(6)[mm$exp], pch=15, cex.lab=1.5, cex.axis=1.5, cex=2)
points(rep(1.7,length(mm$subj))+(mm$exp/10),mm$sd,col=rainbow(6)[mm$exp],pch=16,cex=2)
points(rep(2.7,length(mm$subj))+(mm$exp/10),mm$minimal,col=rainbow(6)[mm$exp],pch=17,cex=2)
axis(side=1,at=c(1,2,3),labels=c("Full","Standardized","Minimal"), cex.axis=1.5)
abline(h=0,lty=2,lwd=1)
dev.off()


```

```{r allExpsRespPlots}

myCol <- c("green","blue","red")
myTitle <- c("Standardize","Full","Minimal")
dt2 <- dt[which(dt$exp == 1),]
for (i in 4:6) { 
  dd <- dt[which(as.integer(dt2$graphType) == i),]
  plot(jitter(dd$effectSize/10),jitter(dd$resp), pch=as.integer(dd$graphType),col=myCol[i-3], xlab="Effect Size",ylab = "Response",yaxt="n",bty="l", main=myTitle[i])
  axis(side=2,at=seq(1,4),labels = seq(1,4))
}



print("% called no or small with full graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==1 & dt$resp < 3)]) / length(dt$Subject[which(as.integer(dt$graphType)==1)]))
  
  
print("% called big with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp == 4)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))

print("% called big  or med with minimal graph:")
print(length(dt$Subject[which(as.integer(dt$graphType)==3 & dt$resp > 2)]) / length(dt$Subject[which(as.integer(dt$graphType)==3)]))
  

```

```{r plotRespByESbyExp}

#Plot Means by Effect Size and Graph Type
myCol <- c("green","blue","red")
myCol <- condCol
myTitle <- c("Full","Standardize","Minimal")
myPCH <- c(16,15,17)
myLWD <- c(2,3,2)
#myLTY <- c(1:3)

#Label positions for each experiment
textLabs <- as.data.frame(matrix(NA,nrow=5,ncol=3))
textLabs[1,] <- c(3.73, 2.73, 1.95)
textLabs[2,] <- c(3.65, 2.8, 1.9)
textLabs[3,] <- c(3.65, 3, 2.15)
textLabs[4,] <- c(3.65, 2.8, 2.25)
textLabs[5,] <- c(3.45, 2.8, 2.2)

xLabs <- as.data.frame(matrix(NA,nrow=5,ncol=4))
xLabs[1,] <- c(.1,.3,.5,.8)
xLabs[2,] <- c(.1,.3,.5,.8)
xLabs[3,] <- c(0,.3,.5,.8)
xLabs[4,] <- c(0,.3,.5,.8)
xLabs[5,] <- c(0,.3,.5,.8)

for (ee in 1:5) { #For each experiment

dt2 <- dt[which(dt$exp == ee),]

#within-subject SEMs
mm <- aggregate(resp ~ effectSize + graphType + Subject, dt2,mean)
m2 <- dcast(mm, Subject ~ graphType + effectSize, value.var = "resp")
m2$subjAvg <- apply(m2[,2:13],1,mean)
m2$adjVal <- mean(m2$subjAvg) - m2$subjAvg
for (i in 2:13) {
  m2[,i+14] <- m2$adjVal + m2[,i]
}

print("check them")
print(round(apply(m2[,2:13],2,mean),2))
print(round(apply(m2[,16:27],2,mean),2))

m3 <- m2[,c(1,16:27)]
colnames(m3)[2:13] <- colnames(m2)[2:13]
m4 <- as.data.frame(apply(m3[,2:13],2,mean),ncol=1,nrow=12)
colnames(m4) <- "resp"
m4$sem <- apply(m3[,2:13],2,sd) / sqrt(length(m3$Subject))
a <- rep(c(1.5,2,3,4),3)
if (ee > 2) { a <- rep(seq(1,4),3)}
m4$x <- a
m4$gt <- c(2,2,2,2,1,1,1,1,3,3,3,3) #1 = full,2 = SD,  3 = small

fn <- paste("Exp",ee,"Slopes.bmp",sep="")
bmp(fn,height=500,width=500)
par(mar=c(5.1,5.1,2.1,.1))
plot(m4$x,m4$resp,col="white",bty="l",xlab="Depicted Effect Size",ylab="Response",xaxt="n",yaxt="n",ylim=c(.99,4.01),cex.lab=1.5)
ggs <- c("Full","SD","Small")
for(i in 1:3) {
  m2 <- m4[which(m4$gt==i),]
  points(m2$x,m2$resp,col=myCol[i],pch=myPCH[i],cex=3)
#  lines(m2$x,m2$resp,lwd=3,col=myCol[i], lty=myLTY[i])
  for (j in 1:4) {
    segments(m2$x[j],m2$resp[j] - m2$sem[j], m2$x[j],m2$resp[j] + m2$sem[j],col=myCol[i])
  }
  
  dd <- dt2[which(dt2$corr < 3 & dt2$graphType == ggs[i]),]
  a <- lm(resp ~ corr, dd)
  a1 <- coefficients(a)[1]
  a2 <- coefficients(a)[2]
  segments(min(m4$x), a1 + a2*min(m4$x), 2, a1 + a2*2, lty=2, col=myCol[i],lwd=myLWD[i])
  
  dd <- dt2[which(dt2$corr > 1.5 & dt2$graphType == ggs[i]),]
  a <- lm(resp ~ corr, dd)
  a1 <- coefficients(a)[1]
  a2 <- coefficients(a)[2]
  segments(2, a1 + a2*2, 4, a1 + a2*4, lty=1, col=myCol[i],lwd=myLWD[i])

  #plot full slope:
#  dd <- dt2[which(dt2$graphType == ggs[i]),]
#  a <- lm(resp ~ corr, dd)
#  a1 <- coefficients(a)[1]
#  a2 <- coefficients(a)[2]
#  segments(min(m4$x), a1 + a2*min(m4$x), 4, a1 + a2*4, lty=3)
  
}
axis(side=1,at=seq(1,4),labels = xLabs[ee,],cex.axis=1.5)
axis(side=2,at=seq(1,4),labels = seq(1,4),cex.axis=1.5)
text(3.3,textLabs[ee,1],"Minimal",adj=c(0,0),cex=1.5, col=myCol[3])
text(3.3,textLabs[ee,2],"Standardized",adj=c(0,0),cex=1.5, col=myCol[2])
text(3.3,textLabs[ee,3],"Full",adj=c(0,0),cex=1.5, col=myCol[1])
dev.off()

} # end for ee exp


```




```{r exploreSDs}

require(reshape2)

ss <- dcast(saveAll, subj + exp ~ graphType, value.var = "coef")
ss$diff12 <- ss$`2` - ss$`1`

sd1 <- sd(saveAll2$coef[which(saveAll2$graphType == 1)])
sd2 <- sd(saveAll2$coef[which(saveAll2$graphType == 2)])
sd3 <- sd(saveAll2$coef[which(saveAll2$graphType == 3)])
rr <- cor(saveAll2$coef[which(saveAll2$graphType == 1)],saveAll2$coef[which(saveAll2$graphType == 2)])
SDwithin <- sqrt(sd1^2 + sd2^2 )

#dZ <- t / sqrt(n)
#dZ <- mDiff / sqrt(SD1^2 + SD2^2 - 2 * SD1 * SD2 * r)


```


