---
title: 'French lexicon project: deciles sampling distributions & individual shift functions'
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

```{r message=FALSE}
# dependencies
library(ggplot2)
library(cowplot)
library(tibble)
# install.packages("devtools")
# devtools::install_github("GRousselet/rogme")
library(rogme)
library(beepr)
source("./functions/skew.txt")
```

```{r}
sessionInfo()
```

# Lexical decision dataset

Data from the [French Lexicon Project](https://sites.google.com/site/frenchlexicon/results).
Click on "French Lexicon Project trial-level results with R scripts.zip".
The `.RData` dataset was created by applying the script `/code/getflprtdata.Rmd`.

```{r}
# get data - tibble = `flp`
load("./data/french_lexicon_project_rt_data.RData")
# columns =
#1 = participant
#2 = rt
#3 = acc = accuracy 0/1
#4 = condition = word/non-word
np <- length(unique(flp$participant))
```

# Compute shift functions for all participants

```{r}
qseq <- seq(0.1,0.9,0.1)
# get data:
flp.s <- subset(flp, condition == "word")
q.w <- matrix(unlist(tapply(flp.s$rt, list(flp.s$participant), quantile, probs = qseq, type = 8)), nrow=9)
flp.s <- subset(flp, condition == "non-word")
q.nw <- matrix(unlist(tapply(flp.s$rt, list(flp.s$participant), quantile, probs = qseq, type = 8)), nrow=9)
q.diff <- q.nw - q.w # non-word - word
```

# Superimpose all shift functions
```{r}
df <- tibble(diff = as.vector(q.diff),
             dec = rep(seq(1,9), np),
             participants = factor(rep(seq(1:np), each = 9)))

df.md <- tibble(diff = apply(q.diff, 1, mean, trim = 0.2),
                dec = seq(1,9))

p <- ggplot(df, aes(x = dec, y = diff, colour = participants)) + theme_classic() +
  geom_line(alpha = 0.5) +
  geom_abline(slope = 0, intercept = 0) +
  geom_line(data = df.md, colour = "black", size = 1) +
  geom_point(data = df.md, colour = "black") +
  scale_colour_viridis_d(option = "B") +
  scale_x_continuous(breaks = seq(1,9)) +
  scale_y_continuous(breaks = seq(-500,700,250)) +
  coord_cartesian(ylim = c(-500, 700)) +
  theme(legend.position = "none",
        plot.title = element_text(size=22),
        axis.title.x = element_text(size = 18),
        axis.text = element_text(size = 16, colour = "black"),
        axis.title.y = element_text(size = 18)) + 
  labs(x = "Deciles", y = "Difference") +
  ggtitle("Non-Word - Word decile differences")
p
p.hsf <- p
```

# 20% trimmed mean group shift function
```{r}
round(apply(q.diff, 1, mean, trim = 0.2))
```

# Median group shift function
```{r}
round(apply(q.diff, 1, median))
```

# Illustrate sampling distributions of decile differences

```{r}
df <- tibble(rt = as.vector(q.diff),
             Decile = factor(rep(seq(1,9), np)))
p <- ggplot(df, aes(x = rt, colour = Decile)) + theme_classic() +
  # geom_density(size = 1) +
  stat_density(geom = "line", position = "identity", size = 1) +
  scale_colour_viridis_d(option = "A", end = 0.9) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  coord_cartesian(xlim = c(-300, 400)) +
  scale_x_continuous(breaks = seq(-300, 400, 100)) +
  theme(plot.title = element_text(size=22),
        axis.title.x = element_text(size = 18),
        axis.text = element_text(size = 16, colour = "black"),
        axis.title.y = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        legend.key.width = unit(1.5,"cm"),
        legend.position = c(0.85,0.7)) +
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  xlab("Difference") +
  ylab("Density") +
  ggtitle("Decile sampling distributions")
p
p.sampdist <- p
```

## Summary figure
```{r, eval = FALSE}
cowplot::plot_grid(p.hsf, p.sampdist,
                          labels = c("A", "B"),
                          ncol = 1,
                          nrow = 2,
                          label_size = 20,
                          hjust = -0.5, 
                          scale=.95)
# save figure
ggsave(filename=('./figures/figure_flp_dec_samp_dist.pdf'),width=8,height=10)
```


## Skewness of sampling distributions
```{r}
apply(q.diff, 1, skew)
```

## Kurtosis of sampling distributions
```{r}
apply(q.diff, 1, kurt)
```

## Stochastic dominance

### Participants with all decile differences > 0
```{r}
pdmt0 <- apply(q.diff > 0, 2, sum)
sum(pdmt0 == 9)
round(100 * sum(pdmt0 == 9) / np, digits = 1)
```
### Participants with all decile differences < 0
```{r}
pdlt0 <- apply(q.diff < 0, 2, sum)
sum(pdlt0 == 9)
round(100 * sum(pdlt0 == 9) / np, digits = 1)
```

## Pattern across deciles

### Spearman function
```{r}
spear<-function(x,y=NULL){
# Compute Spearman's rho
#
if(!is.null(y[1])){
m=elimna(cbind(x,y))
n=nrow(m)
x=m[,1]
y=m[,2]
corv<-cor(rank(x),rank(y))
}
if(is.null(y[1])){
x=elimna(x)
n=nrow(x)
m<-apply(x,2,rank)
corv<-cor(m)
}
test <-corv * sqrt((n - 2)/(1. - corv^2))
sig <- 2 * (1 - pt(abs(test), length(x) - 2))
if(is.null(y[1]))sig<-matrix(sig,ncol=sqrt(length(sig)))
if(sig < 0.05){
  out <- sign(corv)
} else {
  out <- 0
}
out
}
```

### Participants with monotonic increase across deciles
```{r, message=FALSE}
qseq <- seq(0.1,0.9,0.1)
pdmi <- apply(q.diff, 2, spear, qseq)

sum(pdmi == 1)
round(100 * sum(pdmi == 1) / np, digits = 1)
```

### Participants with monotonic decrease across deciles
```{r, message=FALSE}
sum(pdmi == -1)
round(100 * sum(pdmi == -1) / np, digits = 1)
```

# Illustrate sampling distributions of Word deciles

```{r}
df <- tibble(rt = as.vector(q.w),
             dec = factor(rep(seq(1,9), np)))
ggplot(df, aes(x = rt, colour = dec)) + theme_classic() +
  geom_density(size = 1) +
  scale_colour_viridis_d() +
  coord_cartesian(xlim = c(300, 1500)) +
  xlab("Word")
```

### Skewness of sampling distributions
```{r}
apply(q.w, 1, skew)
```

### Kurtosis of sampling distributions
```{r}
apply(q.w, 1, kurt)
```

# Illustrate sampling distributions of Non-Word deciles

```{r}
df <- tibble(rt = as.vector(q.nw),
             dec = factor(rep(seq(1,9), np)))
ggplot(df, aes(x = rt, colour = dec)) + theme_classic() +
  geom_density(size = 1) +
  scale_colour_viridis_d() +
  coord_cartesian(xlim = c(300, 1500)) +
  xlab("Non-Word")
```

### Skewness of sampling distributions
```{r}
apply(q.nw, 1, skew)
```

### Kurtosis of sampling distributions
```{r}
apply(q.nw, 1, kurt)
```


